In this thesis, we study two types of forecasting problems in genetics that require modeling approaches that are both flexible and able to incorporate domain knowledge that may be imperfect and incomplete.

In \cref{ch:doubletrouble}, we develop a Bayesian nonparametric approach for forecasting new and shared variant counts. The approach avoids modeling a possibly finite upper bound while allowing the model to grow as more data come in. The method also has built-in power-law-like growth as a prior while not requiring users to specify a fixed, known power-law rate, but rather learning it from data using an empirical Bayes approach. In \cref{ch:sbirr}, we extend existing Schrödinger bridge approaches to incorporate more flexible forms of domain knowledge, rather than requiring domain scientists to fully specify a model as a reference or prior. In \cref{ch:snapmmd}, we develop a least-squares approach for fitting SDE models that allows model-based forecasting. In applications to, e.g., immune cell activation, our method allows models with neural networks as components while still having biochemistry-informed structure, enabling robust forecasting.

A central methodological theme that emerges from this work is the need to balance flexibility and structure in probabilistic modeling. Flexible models---such as nonparametric or neural approaches---are essential for capturing the complexity of modern scientific data. However, in settings with partial observability and limited signal, flexibility alone is insufficient: models that fit the observed data well may fail to generalize or extrapolate in meaningful ways. Conversely, overly restrictive parametric assumptions may impose incorrect structure and lead to biased or misleading conclusions. The methods developed in this thesis aim to navigate this trade-off by combining flexible modeling components with problem-specific structural assumptions. In the variant discovery setting, this takes the form of Bayesian nonparametric models that capture shared and population-specific variation. In the trajectory inference setting, it involves the use of reference dynamics and iterative projection schemes to constrain the space of plausible trajectories. In the forecasting setting, it requires incorporating structure into SDEs to enable stable and interpretable predictions.

Despite being successful, several challenges still remain that should be addressed in future research. In the variant discovery application, although it is mathematically straightforward to generalize the method to more than two populations, the private-shared structure can become combinatorial---in two populations, there are only three kinds of variants, namely variants private to population 1, private to population 2, and variants that are shared. In a general $p$-population case, there will be $2^p-1$ types. We suspect this brings both statistical and computational challenges. Although we are able to avoid \textit{infinite}-dimensional integrals by leveraging conjugate structure, we still have to solve a numerical integration problem that grows with the number of populations, and this can be challenging in higher dimensions. Last but not least, we only establish power-law bounds rather than exact power-law growth. Future theoretical studies are needed to further understand this class of methods.

In the trajectory inference and forecasting setting, both of our methods in \cref{ch:sbirr} and \cref{ch:snapmmd} require numerically solving SDEs, which is computationally intensive. In the literature, there are methods that avoid simulation, such as \citet{tong2023simulation, zhang2025learning}. However, these methods rely on specific reference families rather than allowing general prior knowledge to be used. The idea used in \citet{bartosh2025sde} of implicitly parameterizing an SDE using marginals and a diffusion term can be a promising approach to avoid simulation, as the implicitly parameterized model can directly predict snapshot observations. Another direction for development is to backtrack, i.e., move time backward. Backtracking in SDEs is not the same as forecasting with inverted time because diffusion processes are not time-symmetric. Ideas in \citet{bartosh2025sde} might also be used by defining the SDE with time horizons before the first observation. Theoretically, when an SDE can be identified using snapshot observations alone is still an open problem, with some progress in gradient fields \citep{lavenant2024toward} and Ornstein-Uhlenbeck processes \citep{guan2024identifying,zhang2025learning}, but not in general. Future theoretical research is needed to understand the fundamental limits of this kind of data.

More broadly, the ideas developed in this thesis are applicable beyond the specific domains considered here. For example, in biodiversity assessments, ideas developed in \cref{ch:doubletrouble} could help predict new species counts across different habitat types using data from a small pilot survey, and ideas developed in \cref{ch:sbirr} and \cref{ch:snapmmd} could be used to infer and forecast individual birds' migration paths from citizen science data that do not identify individual birds. Many problems in science and engineering involve reasoning about partially observed stochastic systems, where the available data provide only indirect information about the underlying processes while useful domain knowledge exists. The perspective of developing methods that balance flexibility and structure provides a general framework for approaching such problems.

In summary, this thesis argues that reliable inference in many genomics applications requires going beyond traditional approaches that focus solely on fitting observed data. Instead, it requires explicitly modeling the latent structures and dynamics that give rise to the data, while balancing flexibility and structure to enable meaningful extrapolation. By developing methods for forecasting unseen structure, reconstructing latent dynamics, and predicting future behavior, this work contributes to a broader understanding of how to perform statistical inference in settings where data are limited, indirect, and inherently incomplete.