New sequencing technologies allowed us to understand various processes in the central dogma: from the genetic architecture of individuals \citep{green2010draft, gravel2011demographic, tennessen2012evolution, lazaridis2014ancient, haak2015massive} to gene expression change over time in single cell level \citep{shapiro2013single, grun2015design}. These different measurements provide different view point of the biological processes of human history, development of cancer, immune system function, etc. Despite large decreases in the cost of sequencing \citep{christensen2015assessing, park2016trends, schwarze2018whole, nawy2014single, ding2022temporal, potter2018single}. Many scientific inquiries could be benefit from having a reliable forecasting method: e.g., forecasting number of unique genetic variants being seen in large, costly follow up samples using a small pilot sample for better data collection design, and forecasting gene expression level change over time in immune cell's life cycles. 

Forecasting is a fundamentally difficult problem that involving making predictions where it is not covered by data, e.g., in new sample sizes or time. Noisiness of genomics data added more complexity to the problem. However biologists often have access to some potentially imperfect domain knowledge that can help the forecasting. A method ideally should make use of such domain knowledge while being flexible to both data and imperfectness of such domain knowledge for better forecasting. 

In this thesis, we focused on two forecasting tasks. The first one naturally arise from large scale population genetics surveys targeting novel (rare) genetic variants \citep{lek2016analysis, karczewski2020mutational}. These variants are fundamentally units of many diseases and potential target for novel therapeutics  \citep{szustakowski2021advancing}. In this set up, a prediction method should accurately estimate the number of novel variants in a new (follow-up) study of multiple populations (e.g., different types of cancer), given data from a typically small, previous (pilot) study, i.e., forecasting beyond the observed \textit{sample sizes}. It is also useful to predict other summaries such as the number of shared or private variants seen in a follow-up studies. In all of these cases, it is crucial to account for novel variants that are shared between groups as opposed to those that are private to a single group.


The second task arises from using single cell sequencing technique to understand gene expression change in cells' life cycle measured by mRNA concentration. One can model the dynamics of mRNA concentration in each cell using a stochastic differential equation \citep[][SDE]{pratapa2020benchmarking}, and there is extensive theory on SDEs from trajectories densely sampled in time. However, scientists cannot measure mRNA concentration continuously in time, but rather only at certain time snapshots. Moreover, since measuring mRNA concentration requires destroying the cell, scientists cannot track the trajectory of any single cell across multiple times. The goal is two fold: 1) reconstructing the distribution of unobserved trajectories and 2) forecast trajectories forward in time. 

Although seemingly distinct, two forecasting tasks face similar challenges. When forecasting number of new variants in population genetics surveys with small amount of pilot samples, the information might not be enough for us to reliably estimate a potential upper bound, meanwhile, when multiple population exist, we need to estimate the correlation structure between the two populations using small amount of samples. When learning SDE from snapshot observations, without further assumptions, there are infinitely many SDEs that can produce the observed snapshots \citep{lavenant2024toward}. 

The lack of information requires methods to be simultaneously flexible to data and having ability to incorporate possibly imperfect and incomplete domain knowledge. In this thesis, we develop a set of probabilistic frameworks for forecasting under such partial observations. 


Concretely, we study two classes of forecasting problems. First, in the context of population genetics, we develop Bayesian nonparametric models for predicting the number and structure of unseen genetic variants across multiple populations, accounting for shared and population-specific variation. The Bayesian nonparametric formalism allowed us to avoid modeling potential upper bound of number of variants being seen thus keeping flexibility while being able to incorporate domain knowledge like usual power-law-like growth of number of variants \citep{Masoero2022}. We adopted an empirical Bayesian-like approach to also use data to inform these domain knowledge. 

Second, in the context of single-cell dynamics, we develop methods for inferring latent stochastic dynamics from snapshot observations and for forecasting their evolution over time. Our methods were able to use a large, flexible class of feasible SDEs (e.g., with nonparametric function class or neural networks) and able to incorporate domain knowledge in a flexible form that do not require domain scientists to encode their knowledge in a single model with all parameters known as in previous methods. Our method is close to empirical Bayes approach that allowed fitting of prior knowledge to observed data. 


Although these problems arise in different domains, they are unified by a common goal: reliable forecasting under limited and indirect observations while solution involving developing probabilistic methods flexible enough to account for data and structured enough to incorporate information from domain knowledge while not requiring domain knowledge to be perfect.
