% \documentclass[ba,preprint]{imsart}% use this for supplement article
\documentclass[ba]{imsart}
%
\pubyear{TBA}
\volume{TBA}
\issue{TBA}
% \doi{0000}
%\arxiv{2207.07020}
\firstpage{1}
\lastpage{1}

\input{ba_header}
\input{shortcuts.tex}
%\externaldocument[supp-]{double_trouble_supplement}
%\usepackage{xr, xr-hyper}



\startlocaldefs
\numberwithin{equation}{section}
\theoremstyle{plain}
\newtheorem{thm}{Theorem}[section]

\endlocaldefs

\begin{document}
\newcommand\blind{1}

\begin{frontmatter}
\title{Double trouble: Predicting new variant counts across two heterogeneous populations}
\runtitle{Double trouble}

\begin{aug}
\author{\fnms{Yunyi} \snm{Shen}\thanksref{addr1}\ead[label=e1]{yshen99@mit.edu}},
\author{\fnms{Lorenzo} \snm{Masoero}\thanksref{addr1}\ead[label=e2]{lom@mit.edu}},
\author{\fnms{Joshua G.} \snm{Schraiber}\thanksref{addr2}\ead[label=e3]{schraibe@usc.edu}},
\and
\author{\fnms{Tamara} \snm{Broderick}\thanksref{addr1}\ead[label=e4]
{tbroderick@mit.edu}}

\runauthor{Y. Shen et al.}

\address[addr1]{Laboratory for Information \& Decision Systems. Massachusetts Institute of Technology.
    \printead*{e1}, % print email address of "e1"
    \printead*{e2}, 
    \printead*{e4}
}

\address[addr2]{
    Department of Quantitative and Computational Biology, USC
    \printead{e3}
    %\printead{u1}
}

%\thankstext{t1}{Some comment}
%\thankstext{t2}{First supporter of the project}
%\thankstext{t3}{Second supporter of the project}

\end{aug}

\begin{abstract}
\input{tex/abstract}
\end{abstract}

\begin{keyword}[class=MSC]
\kwd[Primary ]{62F15}
\kwd[; secondary ]{64G05}
\end{keyword}

\begin{keyword}
\kwd{Bayesian nonparametrics}
\kwd{feature sampling models}
\kwd{beta process}
\kwd{genomic variants} 
\kwd{genomic diversity}
\end{keyword}

\end{frontmatter}

\section{Introduction}
\label{sec:introduction}
\input{tex/intro}

\section{Problem setup}
\label{sec:setup}
\input{tex/setup.tex}

\section{A Bayesian framework for two populations}
\label{sec:bnp_two_pops}
\input{tex/bnp_two_pops.tex}

\section{A natural extension from one population fails}
\label{sec:cant_have_it_all}
\input{tex/cant_have_it_all.tex}

\section{Our model}
\label{sec:our_model} %\label{sec:dependent_prior}
\input{tex/our_model.tex} 

\section{Prediction} 
\label{sec:prediction_all}
\input{tex/prediction_all.tex}

	\subsection{Predicting variant numbers for fixed hyperparameters}
	\label{sec:prediction_fixed}
	\input{tex/prediction_fixed.tex}
	
	\subsection{Empirics for hyperparameters}
	\label{sec:hyperparameter_empirics}
	\input{tex/hyperparameter_empirics.tex}

\section{Power law asymptotics for the number of new variants}
\label{sec:powerlaw}
\input{tex/power_law_main}

\section{Experiments}
\label{sec:experiments}
\input{tex/experiments.tex}

\section{Discussion}
\label{sec:discussion}
\input{tex/discussion}

\begin{acks}[Acknowledgments]
    \input{tex/funding.tex}
    \input{tex/acknowledgements.tex}
\end{acks}
    
%\begin{funding}[Funding]
%    \input{tex/funding.tex}
%\end{funding}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Supplementary Material, if any, should   %%
%% be provided in {supplement} environment  %%
%% with title and short description.        %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{supplement}
\stitle{We here describe the content of supplementary Material for ``Double trouble: Predicting new variant counts across two heterogeneous populations''}
\sdescription{
\internal{
The supplement contains mathematical proofs of the theorems presented in the main text and additional experiments. 
In particular: 
Section S1 contains the proof for our impossibility result (\cref{prop:incomp} in \cref{sec:cant_have_it_all}).
Section S2 contains the proof that our proposed prior satisfies \cref{req:finite_mean,req:inf_mass}.
Section S3 contains the proof that our proposed prior forms a conjugate family to the Bernoulli likelihood process.
Section S4 contains a full derivation of the inference procedure we adopt for single population methods.
Section S5 contains details on the inference procedure we adopt to implement our proposed method.
Section S6 contains details on the setup we adopt for our simulation studies.
Section S7 contains additional results for both simulated and real data.
Section S8 contains proofs of the asymptotic results presented in \cref{sec:powerlaw}.
}}
\end{supplement}
%\begin{supplement}
%\stitle{Title of Supplement B}
%\sdescription{Short description of Supplement B.}
%\end{supplement}

%The following are some other citations to check that the bibliographc style
%file is working correctly: \citet{akaike}, \citet*{akivarsq}, \citet{dyke},
%\citet{greene}, \citet*{kstuart}, \citet{hilbetech}, \citet{hilbe},
%\citet{hilbeglm}, \citet{maddalacntrsq}, and \citet*{companion}.  


\bibliographystyle{ba}
\bibliography{refs}




\end{document}

