% !TEX root = ../double_trouble_main.tex

We next check the quality of our predictions experimentally. First we check that our method performs well on well-specified synthetic data, as well as some particular forms of misspecification. Then we see that our method provides high-quality predictions on a real human cancer genetics dataset and a real human genomic dataset. Throughout, we illustrate the advantages of modeling two populations together rather than treating the two populations separately or all from a single population.

\paragraph{Experimental setup.} In all of our experiments, we will have access to a collection of genetic samples for each of two populations (e.g., two types of cancer). Our validation procedure is reminiscent of cross-validation and analogous to the single population validation procedures of \citet{Masoero2022,  zou2016quantifying}. Within each population, we uniformly at random choose a partition of the data into 20 equally-sized blocks. \minorrevision{Since the dataset size may not be a multiple of 20, block sizes in practice may vary by 1. We will elide  small details of this form in the remainder of this section for clarity of exposition. Also, the population genetics data is substantially larger than our other datasets, so --- to avoid a very large pilot size --- we use 30-fold cross-validation in that case (\cref{sec:real_data}).} 

In each of 20 folds, we reserve one block (unique to this fold) to be the pilot data for its corresponding population, so the final pilot sample has size $\vecsamplesize = (\samplesize{1}, \samplesize{2})$. The remaining data will serve as potential follow-up data. Each method has access only to the pilot data (across both populations), and we use the follow-up data (across both populations) to evaluate the accuracy of each method. None of the methods we consider in this paper use ordering information on the pilot data, but to generate all of our figures, we choose an ordering on the pilot data uniformly at random. Likewise, we choose an ordering on the follow-up data uniformly at random. In our figures, we present first the pilot data from population 1, then the pilot data from population 2, then the follow-up data from population 1, then the follow-up data from population 2. We can think of these figures as encapsulating a series of evaluations as follows. We train every method on the full pilot data in populations 1 and 2. Then we progressively test on follow-up sample sizes $\vecfuturesamplesize = (\futuresamplesize{1},\futuresamplesize{2})$ of the following forms: $(1,0), (2,0), \ldots, (19 \samplesize{1},0), (19 \samplesize{1},1), (19 \samplesize{1},2), \ldots, (19 \samplesize{1}, 19 \samplesize{2})$. In this way, we capture a wide range of ratios between follow-up sample sizes across the two populations: from all follow-up data in a single population to a ratio of follow-up sizes that matches the ratio in the pilot.
To summarize our results across all 20 folds, we will show an empirical mean across folds, both for each prediction method as well as for the ground truth. And we will show a shaded one-standard deviation interval for the prediction methods. The ground truth showed very small variation across folds so the interval appears almost invisible. %\tb{ultimately we will say something about ground truth here too once the results are in.} %\tb{Are the intervals for the ground truth very small? can we say something here?}

Recall that we are also interested in predicting the number of $\vecoccurrence$-tons in a follow-up sample of size $\vecfuturesamplesize$ (\cref{eq:news_freq}); that is, we are interested in predicting the number of new variants appearing exactly $\vecoccurrence$ times across the two populations. To evaluate our prediction in this case, we consider just the extreme where we use all possible data in the follow-up. For each value of $\vecoccurrence$ with $\occurrence{1}, \occurrence{2}$ each less than a threshold, we compute the \emph{relative residual}: namely, the signed difference between a method's prediction and the observed value in the reserved follow-up data, normalized by the observed value. We plot an empirical average of this quantity across the folds. We also report the actual observed values and absolute (un-normalized) residuals in Section S7.3 in the Supplementary Material \citep{supplement}, again averaged across the folds.

\paragraph{Comparisons.} To the best of our knowledge, there are no alternative methods for the prediction task we are interested in here. However, there are a number of methods that predict the number of variants in a follow-up given pilot data in a single population~\citep{Masoero2022, chakraborty2019using,gravel2014predicting}. So we consider two natural competitors in our experiments. First, we define the \emph{independent} two-population \revision{application} of a single-population method as follows. Given $\vecsamplesize$ pilot samples and $\vecfuturesamplesize$ follow-up samples, we train the single-population method first on the $\samplesize{1}$ pilot samples from population 1 and predict the number of variants expected in $\futuresamplesize{1}$ follow-up samples from population 1. We separately train the single-population method on the $\samplesize{2}$ samples from population 2 and predict the number of variants expected in $\futuresamplesize{2}$ follow-up samples from population 2. We report the sum of the predicted number of variants across both populations as the final prediction. 
In general, we expect the independent method to overcount the number of variants when there are follow-up samples from both populations. We do not include the independent method below when we compare performance in predicting $\vecoccurrence$-tons. In particular, whenever both elements of $\vecoccurrence$ are strictly positive, the independent method predicts there are zero $\vecoccurrence$-tons. We provide the \revision{prediction, its supporting theory, and how we perform hyperparameter selection} in Section S4.2 of the Supplementary Material of \citep{supplement}. In the real data we analyze at the sizes we consider, this zero prediction is unrealistic; see Figures S8 and S9 in the Supplementary Material \citep{supplement} for the ground-truth real-data $\vecoccurrence$-ton values. %(Also, see \cref{supp-fig:proposed_kton,supp-fig:i3BP_kton,supp-fig:d3BP_kton} for the ground-truth synthetic $\vecoccurrence$-ton values.)

%The independent method essentially assumes each variant is unique to a single population. From that perspective, when $\vecoccurrence = (\occurrence{1},0)$, we can take the population-1 prediction for the number of $\occurrence{1}$-tons and use it as the independent-method prediction for the number of $\vecoccurrence$-tons; an analogous prediction holds when $\occurrence{1} = 0$ and $\occurrence{2} > 0$. Whenever both elements of $\vecoccurrence$ are strictly positive, the independent method predicts there are zero $\vecoccurrence$-tons. In the real data we analyze at the sizes we consider, this zero prediction is unrealistic (see~\Cref{supp-fig:proposed_kton},~\ref{supp-fig:i3BP_kton},~\ref{supp-fig:d3BP_kton} for raw number of $\vecoccurrence$-tons in synthetic experiments and~\Cref{supp-fig:brca_luad_kton},~\ref{supp-fig:seu_bgr_kton} in real data experiments). 

Second, we define the \emph{dependent} two-population \revision{application} of a single-population method. In this case, we train the single-population method on all of the $\samplesize{1} +\samplesize{2}$ pilot samples, as though they are the pilot samples from a single population. To form a prediction on a follow-up sample of size $\vecfuturesamplesize$, we report the prediction on a single-population follow-up of size $\futuresamplesize{1} + \futuresamplesize{2}$. 
In general, we expect the dependent method to struggle when both (a) the two populations exhibit notably different variant growth rates and (b) the ratio of populations in the follow-up does not closely mirror the pilot \revision{because this mismatch exacerbates prediction errors. For example, if one population contains many rare variants while the other does not, changing the sampling proportions between pilot and follow-up will substantially impact variant discovery rates in ways the homogeneous model cannot account for.}
To predict the number of $\vecoccurrence$-tons, we predict the number of variants, conditional on the pilot, that we would expect to see $\occurrence{1}$ times out of $\futuresamplesize{1}$ follow-up samples and also $\occurrence{2}$ times out of $\futuresamplesize{2}$ distinct follow-up samples (all from the same assumed single population). For single-population methods that do not provide $\occurrence{}$-ton predictions (for scalar $\occurrence{}$), we will therefore not be able to provide $\vecoccurrence$-ton predictions in the dependent \revision{application}. 
However, we are able to provide $\vecoccurrence$-ton predictions for the special case of the single-population 3BP method due to \citet{Masoero2022}.
\internal{In particular, in Section S4.1 of the Supplementary Material \citep{supplement}, we describe how we modify \citet{Masoero2022} to predict vector-valued $\vecoccurrence$-tons; we provide theory analogous to \cref{prop:predict_kr_tons} and describe the method we employ to fit the hyperparameters}.
%\internal{see Section S4.1 of the Supplementary Material \citep{supplement} for a discussion on modifying \citet{Masoero2022} to predict vector values $\vecoccurrence$-tons, including supporting theory and the method employed to fit the hyperparameters}.

Since there are many single-population methods, in general we will select just one single-population method to use when reporting the independent and dependent results. In each case, we will try to use the single-population method that is best in a relevant sense, which we make precise in each case below.


\subsection{Synthetic experiments}
\label{sec:simulation}
\input{tex/simulation.tex}

\subsection{Real data}
\label{sec:real_data}
\input{tex/real_data.tex}