Practitioners are often interested in the possible paths taken by a population of particles moving from one location to another in a given space. For example, biologists want to understand how gene expression, as measured by mRNA levels, changes when normal cells transform into cancer cells. A deeper understanding might aid development of methods to prevent or treat cancer. One can model the dynamics of mRNA concentration in each cell using a stochastic differential equation (SDE), and there is extensive theory on SDEs from trajectories densely sampled in time. However, scientists cannot measure mRNA concentration continuously in time, but rather only at certain time snapshots. Moreover, since measuring mRNA concentration requires destroying the cell, scientists cannot track the trajectory of one cell across multiple times. 

Recent work has demonstrated the potential of using Schr\"odinger bridges (SBs) to infer possible trajectories connecting two time snapshots \citep*{pavon2021data,de2021diffusion, Vargas2021,koshizuka2022neural, Wang2023}. This approach learns a pair of forward-backward SDEs that can transport particles between two time points (forward or backward, respectively). One can then use the learned SDEs to numerically sample from the latent population of trajectories. 
To extend these methods to data with multiple time snapshots, 
one can apply vanilla SBs separately over each pair of consecutive time points.  
%\revised{But then the connected bridges exhibit kinks in the form of sharp turns in trajectories at connection time points}, 
But the learned dynamics may fail to capture long-term dependencies, seasonal patterns, and cyclic behaviors.
\citet{lavenant2024toward} show that other natural extensions to multiple time steps \citep{chen2019multi, lavenant2024toward} still decompose into vanilla SBs between adjacent time steps. 

To overcome these issues, \citet{chen2024deep} proposed to instead use the SDEs in SBs to govern particles' velocity, instead of location. As with the methods above, their method allows the dynamics to pass through multiple time snapshots.
In addition, since their method generates smooth trajectories, trajectory information can be shared across time intervals. 

A practical issue with all methods above, including \citet{chen2024deep}, though, is that they require a single pre-defined reference measure. In particular, these methods minimize relative entropy between the process and the reference. 
Ideally, the reference dynamics should be close to the true, latent dynamics. While scientists often have some prior knowledge of how their system works \citep[e.g., the parametric form of the underlying dynamical model, as in][]{pratapa2020benchmarking}, it is rare that they can access all parameter values. Consequently, practitioners often default to using Brownian motion as the reference dynamic. In fact, \citet{schrodinger1932theorie} originally used Brownian motion to model particles moving under thermal fluctuations in a closed system; for a historical review, see \citet{leonard2013survey}. In this context, solving the SB problem is equivalent to solving an entropy-regularized optimal transport problem \citep{cuturi2013sinkhorn}. However, this approach may not be ideal for many open systems with energy intake. For instance, biological systems are such systems and do not necessarily adhere to the same entropy laws as gas molecules under thermal fluctuations. \citet[][Chapter 6]{schrodinger1946life} himself  famously remarked ``What organisms feed upon is \textit{negative} entropy.''

Methods beyond the SB literature face similar (and additional) limitations. 
For instance, TrajectoryNet \citep{tong2020trajectorynet} combines continuous normalizing flows with a soft constraint based on dynamic optimal transport. But it can be interpreted as an SB method with a single reference;
see \citet[sections 3.2 and 4.1]{tong2020trajectorynet}. See \cref{app:related-work} for more related work. 


Due to the limitations of existing methods, we propose a new approach that (1) infers the distribution of trajectories from sample snapshots at multiple time points and (2) requires the user to specify only a parametric family of reference dynamics rather than a single one. Our method iterates between two steps. (i) Given our current best guess of the latent dynamics, we learn a piecewise SB and sample the resulting trajectories. (ii) We use the learned SB to refine our best guess of the latent dynamics within the reference family. 
Since the estimate of latent dynamics uses information from all times, we expect our method to share information across time intervals.
In our experiments (\cref{ch3sec:experiments-main}), we find that our method is more accurate than both vanilla SBs and alternatives that nontrivially handle multiple time points. %Moreover, our method is substantially faster than the latter methods in all experiments, sometimes over 40 times faster. 
Moreover, our method is faster than the latter methods in all experiments.\footnote{Code to reproduce the experiments is available at \href{https://github.com/YunyiShen/SB-Iterative-Reference-Refinement}{https://github.com/YunyiShen/SB-Iterative-Reference-Refinement}.} %, sometimes over 40 times faster. 



In concurrent work, \citet{zhang2024joint} independently proposes an iterative procedure that allows for a family of reference dynamics. While in our work we use SBs to recover continuous-time evolution of the underlying dynamics, \citet{zhang2024joint} uses an entropic optimal transport framework and aims to recover the underlying dynamics solely at points with observations. And while we allow a general reference family, \citet{zhang2024joint} uses a linear (Ornstein-Uhlenbeck) reference family. We share the same  motivation though: moving away from a fixed reference dynamic to improve trajectory inference.


\subsection*{Additional related works}
In this section, we expand our related work discussion beyond the Schrödinger bridges literature to works that are closely related to our problem.

\textbf{Generative modeling.} Recently, there has been significant progress in the literature on modeling flows or transports between probability distributions. These approaches are built on a variety of frameworks, such as score-based generative modeling \citep{song2019generative, song2020score}, diffusion models \citep{ho2020denoising}, Schrödinger bridges \citep{de2021diffusion,pavon2021data,Vargas2021,Wang2023}, continuous normalizing flows \citep{chen2018neural,grathwohl2018ffjord}, and flow matching \citep{lipman2022flow}. However, the majority of these works focus on generative modeling, where the goal is to transform a noise distribution, such as a Gaussian, into a data distribution to generate samples. Moreover, they typically address transport between distributions at only two time points within one being typically Gaussian. In contrast, our focus is not on learning how to generate data from noise, but on learning and reconstructing trajectories that capture the underlying system dynamics as they evolve across multiple time steps. 

\textbf{Trajectory inference.} Recent studies have been focused on interpolating the trajectories of individual samples at multiple time points with optimal  transport \citep{schiebinger2019optimal, yang2018scalable}. However in certain situations just mere interpolation is not enough, since the learned dynamics may fail to capture long-term dependencies, seasonal patterns, and cyclic behaviors. Other works in the optimal transport literature helped to overcome this issue: \citet{hashimoto2016learning} proposed to reconstruct trajectories using a particular regularized recurrent neural network (RNN) with Wasserstein gradient flow loss. \citet{bunne2022proximal} expanded this approach by modeling the evolution of the trajectories of interest using the Jordan-Kinderlehrer-Otto (JKO) scheme. In particular, the authors developed \textit{JKOnet}, a neural architecture that can compute the JKO flow, from which the energy landscape that governs the population's dynamics can be learnt. \textit{TrajectoryNet} by \citet{tong2020trajectorynet} combines ideas from dynamic optimal transport with continuous normalizing flows (CNFs). In particular, the authors use a CNF to generate continuous-time non-linear sample trajectories from multiple time points. And they also propose a regularization that encourages a straight trajectory based on dynamic optimal transport theory. \citet{huguet2022manifold} recently proposed another flow-based method,
Manifold Interpolation Optimal-transport Flow (\textit{MIOFlow}). The main idea of this work is to solve the flow problem in a manifold embedding. The authors use a neural ODE \citep{chen2018neural} to transport a sampling of high dimensional data points between time points such that (1) the transport occurs on a manifold defined by samples, (2) the transport is penalized
to agree with measured time points using Wasserstein, and (3) the transport is stochastic.

All of these methods can be viewed as optimizing an entropy-regularized (unbalanced) optimal transport loss between observed and simulated samples, requiring backpropagation through optimal transport solvers \citep{cuturi2013sinkhorn, cuturi2022optimal}. Essentially, these methods solve a Schrödinger bridge problem, each utilizing a specific reference measure based on the chosen method. For example, \citet{tong2020trajectorynet} simulates samples using CNFs with a specific dynamic optimal transport-induced regularization, and this leads to approximately solving a Schrödinger bridge problem where the volatility term is 0 and the reference is a Brownian motion. Therefore, similarly to what we discussed in the main text for \citet{chen2024deep}, all these methods suffer from the fact that it is as if they need a pre-defined reference measure. And it is very rare that scientists have access to all the information about the correct reference measure for their problems. Our proposed method directly addresses this challenge by moving beyond a fixed reference process to improve trajectory inference. We emphasize that our main contribution lies in the reference refinement step, which, in theory, could be integrated with the methods discussed here as well as Schrödinger bridge methods outlined in the main text. This is an exciting direction for future work.


